---
date: "2023-06-13"
output:
  html_document:
    theme: united
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE)
```

# Comparison between nanoCT results using "peaks" or "bins" approach
This vignette compares the final annotations of the same nanoCT data analysed with [peaks](https://fansalon.github.io/vignette_single-cell-nanoCT.html) or [bins](https://fansalon.github.io/vignette_single-cell-nanoCT_bins.html).\

To compare the results the cells (barcodes) associated to a given cell-type in the "bin" nanoCT results were matched with the respective cell-types in the "peaks" nanoCT results. In other words, we counted how many of the X cells annotated as 'Inhibithory Neurons 1' in the "bin" nanoCT results were annotated as 'Inhibithory Neurons 1' in the "peaks" nanoCT results, how many as 'Excitatory Neurons 1', etc.


## Load libraries
```{r load_lib}
library(ggplot2)
library(reshape2)
```
\


## Load the data
```{r load_data}

# Read RDS
bins <- readRDS("~/NatProt/nanoscope/nanoscope_final_bins_5kb.rds")
peaks <- readRDS("~/NatProt/nanoscope/nanoscope_final_peaks.rds")
```
\

## Count number of cells for cell-type in bins and peaks, for all the modalities
```{r count}

### bins
# for each modality, get barcode and annotation
bins.ls <- lapply(bins, function(x){
  data.frame(barcode=rownames(x@meta.data),
             bins=as.character(x@meta.data[,"layer2_annotation"]))})

### peaks
# for each modality, get barcode and annotation
peaks.ls <- lapply(peaks, function(x){
  data.frame(barcode=rownames(x@meta.data),
             peaks=as.character(x@meta.data[,"layer2_annotation"]))})

### merge
merged.ls <- list()
mat.ls <- list()
for (mod in names(bins.ls)) {
  merged.ls[[mod]] <- merge(bins.ls[[mod]],peaks.ls[[mod]],by="barcode",all=T,sort=F)
  merged.ls[[mod]][is.na(merged.ls[[mod]])] <- "N/A"
  # convert to matrix
  mat.ls[[mod]] = melt(as.matrix(table(merged.ls[[mod]]$peaks, merged.ls[[mod]]$bins)))
}



```
\

## Plot
```{r, fig.width = 8, fig.height = 7}

# ATAC
ggplot(data = mat.ls$ATAC_TATAGCCT, aes(x=Var1, y=Var2, fill=log10(value+1))) + 
  geom_tile() +
  geom_text(aes(label=value),color="white") +
  # x-axis
  theme(axis.text.x=element_text(angle=45, hjust=1, size = 10)) + 
  theme(axis.title.x = element_text(size= 14)) +
  xlab("peaks") +
  theme(axis.title.x = element_text(size= 14)) +
  # y-axis
  ylab("bins") +
  theme(axis.text.y=element_text(size = 10)) +
  theme(axis.title.y = element_text(size= 14)) +
  ggtitle("ATAC") +
  theme(plot.title = element_text(size=14,hjust=0.5,face='bold')) + 
  guides(fill=guide_legend(title="Number of cells\nlog10 + 1"))

# H3K27ac
ggplot(data = mat.ls$H3K27ac_ATAGAGGC, aes(x=Var1, y=Var2, fill=log10(value+1))) + 
  geom_tile() +
  geom_text(aes(label=value),color="white") +
  # x-axis
  theme(axis.text.x=element_text(angle=45, hjust=1, size = 10)) + 
  theme(axis.title.x = element_text(size= 14)) +
  xlab("peaks") +
  theme(axis.title.x = element_text(size= 14)) +
  # y-axis
  ylab("bins") +
  theme(axis.text.y=element_text(size = 10)) +
  theme(axis.title.y = element_text(size= 14)) +
  ggtitle("H3K27ac") +
  theme(plot.title = element_text(size=14,hjust=0.5,face='bold')) + 
  guides(fill=guide_legend(title="Number of cells\nlog10 + 1"))

# H3K27me3
ggplot(data = mat.ls$H3K27me3_CCTATCCT, aes(x=Var1, y=Var2, fill=log10(value+1))) + 
  geom_tile() +
  geom_text(aes(label=value),color="white") +
  # x-axis
  theme(axis.text.x=element_text(angle=45, hjust=1, size = 10)) + 
  theme(axis.title.x = element_text(size= 14)) +
  xlab("peaks") +
  theme(axis.title.x = element_text(size= 14)) +
  # y-axis
  ylab("bins") +
  theme(axis.text.y=element_text(size = 10)) +
  theme(axis.title.y = element_text(size= 14)) +
  ggtitle("H3K27me3") +
  theme(plot.title = element_text(size=14,hjust=0.5,face='bold')) + 
  guides(fill=guide_legend(title="Number of cells\nlog10 + 1"))


```